---
title: "2Process Data"
output: html_notebook
---

This script assumes you have run `1_Download_Data.R` 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(rgeos)
library(wiqid)
memory.limit(memory.limit() *2^30) #on Windows computer
sessionInfo()
```


##1. Load Florida Tracheophytes occurrences
```{r, load Tracheophyte data}
load("./Raw Data/Tracheophytes with Coordinates/tracheophytes_coords.RData")
load("./Raw Data/Tracheophytes without Coordinates/tracheophytes_nocoords.RData")
load("./florida_state.RData")
load("./florida.RData")

```

Some ancillary functions that will be used later in the script: 

```{r, ancillary functions}
renameCounties <- function(df) {


df$county <- gsub(df$county, pattern='(', replacement='', fixed=T)
df$county <- gsub(df$county, pattern=')', replacement='', fixed=T)
df$county <- gsub(df$county, pattern=' County', replacement='', fixed=T)
df$county <- gsub(df$county, pattern='County of ', replacement='')
df$county <- gsub(df$county, pattern=' Cty.', replacement='')
df$county <- gsub(df$county, pattern=' Cty', replacement='')
df$county <- gsub(df$county, pattern=' Co.', replacement='')
df$county <- gsub(df$county, pattern='Ste ', replacement='Saint ')
df$county <- gsub(df$county, pattern='St. ', replacement='Saint ')
df$county <- gsub(df$county, pattern='St ', replacement='Saint ')
df$county <- gsub(df$county, pattern='[', replacement='', fixed=T)
df$county <- gsub(df$county, pattern=']', replacement='', fixed=T)
df$county <- gsub(df$county, pattern=' FL', replacement='')

df$county <- tolower(df$county)
df$county <- capwords(df$county)

df$county <- gsub(df$county, pattern='Dade', replacement='Miami-Dade', fixed=T)
df$county <- gsub(df$county, pattern='Miami-dade', replacement='Miami-Dade', fixed=T)
df$county <- gsub(df$county, pattern='Union County', replacement='Union', fixed=T)
df$county <- gsub(df$county, pattern='Hillsboro', replacement='Hillsborough', fixed=T)
df$county <- gsub(df$county, pattern='De Soto', replacement='Desoto', fixed=T)
df$county <- gsub(df$county, pattern='Walto', replacement='Walton', fixed=T)
df$county <- gsub(df$county, pattern='Monre', replacement='Monroe', fixed=T)
df$county <- gsub(df$county, pattern='Liberity', replacement='Liberty', fixed=T)
df$county <- gsub(df$county, pattern='Hillsborongh', replacement='Hillsborough', fixed=T)
df$county <- gsub(df$county, pattern='Waukulla', replacement='Wakulla', fixed=T)
df$county <- gsub(df$county, pattern='Alachuca', replacement='Alachua', fixed=T)
df$county <- gsub(df$county, pattern='Wahulla', replacement='Wakulla', fixed=T)
df$county <- gsub(df$county, pattern='Semiole', replacement='Seminole', fixed=T)
df$county <- gsub(df$county, pattern='Hillsboroughugh', replacement='Hillsborough', fixed=T)
df$county <- gsub(df$county, pattern='Waltonn', replacement='Walton', fixed=T)
df$county <- gsub(df$county, pattern='Hillsboroughngh', replacement='Hillsborough', fixed=T)
df$county <- gsub(df$county, pattern='Alachue', replacement='Alachua', fixed=T)
df$county <- gsub(df$county, pattern='Miami Miami-Dade', replacement='Miami-Dade', fixed=T)
df$county <- gsub(df$county, pattern='Gadsen', replacement='Gadsden', fixed=T)
df$county <- gsub(df$county, pattern='Gasden', replacement='Gadsden', fixed=T)
df$county <- gsub(df$county, pattern='Wapulla', replacement='Wakulla', fixed=T)
df$county <- gsub(df$county, pattern='Duvall', replacement='Duval', fixed=T)
df$county <- gsub(df$county, pattern='Escambra', replacement='Escambia', fixed=T)
df$county <- gsub(df$county, pattern='Florida Lake', replacement='Lake', fixed=T)
df$county <- gsub(df$county, pattern='Florida Gulf', replacement='Gulf', fixed=T)
df$county <- gsub(df$county, pattern='Colombia', replacement='Columbia', fixed=T)
df$county <- gsub(df$county, pattern='Wakullla', replacement='Wakulla', fixed=T)
df$county <- gsub(df$county, pattern='Oskaloosa', replacement='Okaloosa', fixed=T)
df$county <- gsub(df$county, pattern=',', replacement='', fixed=T)
df$county <- gsub(df$county, pattern='Miami - Miami-Dade', replacement='Miami-Dade', fixed=T)
df$county <- gsub(df$county, pattern='Calhoune,', replacement='Calhoun', fixed=T)
df$county <- gsub(df$county, pattern='Suwanee,', replacement='Suwannee', fixed=T)
df$county <- gsub(df$county, pattern='Poek', replacement='Polk', fixed=T)
df$county <- gsub(df$county, pattern='Duvalole', replacement='Duval', fixed=T)
df$county <- gsub(df$county, pattern='Liverty', replacement='Liberty', fixed=T)
df$county <- gsub(df$county, pattern='Saint John', replacement='Saint Johns', fixed=T)
df$county <- gsub(df$county, pattern='Saint Johnss', replacement='Saint Johns', fixed=T)
df$county <- gsub(df$county, pattern='Franklin Florida United States', replacement='Franklin', fixed=T)
df$county <- gsub(df$county, pattern='Miami', replacement='Miami-Dade', fixed=T)
df$county <- gsub(df$county, pattern='Miami-Dade-Dade', replacement='Miami-Dade', fixed=T)
df$county <- gsub(df$county, pattern='Miami-Dade-Dade', replacement='Miami-Dade', fixed=T)
df$county <- gsub(df$county, pattern='Calhoune', replacement='Calhoun', fixed=T)
df$county <- gsub(df$county, pattern='Suwanee', replacement='Suwannee', fixed=T)
df$county <- gsub(df$county, pattern='Franklyn', replacement='Franklin', fixed=T)
df$county <- gsub(df$county, pattern='Martin County', replacement = 'Martin', fixed=T)
return(df)
}


compare <- function(this, that) {


issues <- NULL

for (i in 1:length(this)){
  test<-agrep(this[i], that, ignore.case = TRUE)
  if (length(test)>1) {
    issues <- c(issues, as.character(this[i]))
  }
}

return(issues)
}

printIssues <- function(issues, compareList) {
for (i in 1:length(issues)) {
print(agrep(issues[i], compareList, value=T, ignore.case = TRUE))
print("...")
}
}

```
###  Records with coordinates : 
    - Clip to Florida state polygon 
    - Assign 'geo_county' as the county the point falls in 
```{r, process records with coordinates, cache=T}
#Assign coordinates and projection: 
coordinates(tracheophytes_coords) <- c(134, 133)
proj4string(tracheophytes_coords) <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0") 
proj4string(florida_state) <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

#Select only those records that fall within the state polygon 
tracheophytes_coords <- tracheophytes_coords[florida_state,]

#Assign counties to each of the records 

#First, remove extra levels of 'county':
tracheophytes_coords@data$county <- factor(tracheophytes_coords@data$county)


#create a column called geo_county for storing the county of each record as determined by overlaying county level polygons 
tracheophytes_coords[["geo_county"]]<-rep("NA", dim(tracheophytes_coords)[1])

#Remove extraneous counties from florida shapefile
florida@data$NAME_2 <- factor(florida@data$NAME_2)
counties<-unique(florida@data$NAME_2) #There are 67 counties


tracheophytes_coords@data$geo_county <- as.character(over(tracheophytes_coords, florida)$NAME_2)

temp <- tracheophytes_coords@data
nrow(temp)
temp <- subset(temp, temp$county != "")
nrow(temp)

temp2 <-subset(temp, temp$county != temp$geo_county)
temp3 <- temp2[, c("gbifID", "county", "geo_county")]
write.csv(temp3, file="./Intermediate Data/check_coordinate_data.csv")

flagged_coordinate_data <- read.csv("./Intermediate Data/flagged_coordinate_data.csv")

temp <- subset(flagged_coordinate_data, flagged_coordinate_data$FLAG == "FLAG")

to_discard <- temp$gbifID

tracheophytes_coords <- subset(tracheophytes_coords, !(tracheophytes_coords$gbifID %in% to_discard))

temp2 <- subset(tracheophytes_coords, tracheophytes_coords$county == "")

write.csv(temp2, "./Intermediate Data/records_with_no_county.csv")
temp3 <- subset(temp2, temp2$family == "Anacardiaceae")
nrow(temp3)

#assumption: don't worry about citizen scientist records that are missing a county field
temp4 <- subset(temp3, temp3$basisOfRecord != "HUMAN_OBSERVATION")
table(temp4$species)
write.csv(temp4, "./Intermediate Data/anacardiaceae_records_with_no_county.csv")

#There are now 148981 records of Tracheophytes with coordinates in the state of Florida that have been checked to discard records where 
# the geographically assigned county doesn't match the county field. 

save(tracheophytes_coords, file="./Intermediate Data/cleaned_tracheophytes_coords.RData")


```
    

### Records without coordinates: 
    - Clean county names
    - Remove records from counties not in Florida or uncertain
    - Assign 'geo_county' as the name of the county 
    
```{r, process records without coordinates, cache=T} 

#Remove those records without state or county
tracheophytes_nocoords <- subset(tracheophytes_nocoords, tracheophytes_nocoords$stateProvince != "")
tracheophytes_nocoords <- subset(tracheophytes_nocoords, tracheophytes_nocoords$county != "")
tracheophytes_nocoords <- subset(tracheophytes_nocoords, tracheophytes_nocoords$countryCode == "US")
tracheophytes_nocoords$stateProvince <- factor(tracheophytes_nocoords$stateProvince) 

unique(tracheophytes_nocoords$stateProvince)
#There are two levels, florida and Florida 

tracheophytes_nocoords$stateProvince <- "Florida"


#rename counties 
tracheophytes_nocoords <- renameCounties(tracheophytes_nocoords)
countiesinList <- unique(tracheophytes_nocoords$county)

#Check for additional county misnaming 
issues2<- compare(countiesinList, countiesinList)
printIssues(issues2, countiesinList)

#Now only keep those records that are in legitimate counties 
tracheophytes_nocoords <- subset(tracheophytes_nocoords, tracheophytes_nocoords$county %in% counties)

#There are 255, 126 county-level records of Tracheophytes in Florida 


#So the dimensions match that of tracheophyte_coords
tracheophytes_nocoords[["geo_county"]]<-tracheophytes_nocoords$county

save(tracheophytes_nocoords, file = "./Intermediate Data/cleaned_nocoords_data.RData")
```
### rbind records with coordinates and records without coordinates together
```{r, combine records with and without coordinates back together, cache=T}

tracheophytes_messyCollectors <- rbind(as.data.frame(tracheophytes_coords), tracheophytes_nocoords)
rm(tracheophytes_coords)
rm(tracheophytes_nocoords)

save(tracheophytes_messyCollectors, file="./Intermediate Data/tracheophytes_messyCollectors.RData")
```

## 2a. Clean collector names using openRefine
```{r use openRefine}
load("./Intermediate Data/tracheophytes_messyCollectors.RData")

listCollectors <- tracheophytes_messyCollectors[, c(1,71)]
write.csv(listCollectors, file = "./Intermediate Data/listCollectors.csv")
#Process data in openRefine the read in the file with the corrections: 


listCollectors_refined <- read.csv("./Intermediate Data/listCollectors_refined.csv")
#Because this was initially cleaned before some of the records were removed for having counties that don't match the geo_county, we need
# to remove some of these: 

IDs <- tracheophytes_messyCollectors$gbifID

listCollectors_refined2 <- subset(listCollectors_refined, listCollectors_refined$gbifID %in% IDs)

tracheophytes <- tracheophytes_messyCollectors

tracheophytes$correctedCollector <- listCollectors_refined2$recordedBy


```



#3 Fix incorrect dates 

```{r fix incorrect dates}



#Start at tail ends of years, and work towards middle until find verified record that is the correct year. Goal: eliminate mistakes on either end. Currently not worrying about mistakes that still fall within lifespan of collector. 

#Walter M. Buswell (professor at UMiami Dept of Botany in 40s)
#His only Schinus record is from 1936. Also has a herbarium named after him. 

#Robert Kenneth Godfrey (1911-2000). Also has a herbarium named after him.

#Robert Kral was born in 1926


#Daniel Bertram Ward (1928-2016)

#K. A. Bradley  == Keith Bradley Alive

#J. Beckner 

#The issues that I found are in the csv file incoorectYearsCollectors.csv

fixes <- read.csv("./Intermediate Data/incorrectYearsCollectors.csv")


#Some other issues: There are some records that are NA for year and or collector
#The two records of Schinus that have no year don't have years on their specimen sheet

#Look at each occurrence record for Schinus to spot irregularities in collector or year or county 





```

# Implement corrections:
```{r implement corrections, cache=T}


tracheophytesClean <- tracheophytes

tracheophytesClean$correctedYear <- tracheophytesClean$year
                               



#have to remove fixes for observations that are no longer in the data set 

fixes2 <-subset(fixes, fixes$GBIF_ID %in% IDs)


fixes2flags <- subset(fixes2, fixes2$CorrectedCollector == "FLAG" | fixes2$CorrectedYear == "FLAG")
fixes2_tofix <- subset(fixes2, fixes2$CorrectedCollector != "FLAG" & fixes2$CorrectedYear != "FLAG")
fixes2_tofix_years <- subset(fixes2_tofix, fixes2_tofix$CorrectedYear != "")
fixes2_tofix_names <- subset(fixes2_tofix, fixes2_tofix$CorrectedCollector != "")

for (i in 1:nrow(fixes2_tofix_years)) {
  tracheophytesClean[which(tracheophytesClean$gbifID == fixes2_tofix_years$GBIF_ID[i]),]$correctedYear <- as.character(fixes2_tofix_years$CorrectedYear[i])
}



for (i in 1:nrow(fixes2_tofix_names)) {
  tracheophytesClean[which(tracheophytesClean$gbifID == fixes2_tofix_names$GBIF_ID[i]),]$correctedCollector <- as.character(fixes2_tofix_names$CorrectedCollector[i])
}


#remove records where the year or collector are flagged: 

toremove <- c(which(tracheophytesClean$correctedYear == "FLAG"), which(tracheophytesClean$correctedCollector == "FLAG"))

tracheophytesClean <- subset(tracheophytesClean, !(tracheophytesClean$gbifID %in% toremove))

toremove <- c(which(tracheophytesClean$correctedYear == "FLAG"), which(tracheophytesClean$correctedCollector == "FLAG"))


tracheophytesClean <- subset(tracheophytesClean, !(tracheophytesClean$gbifID %in% toremove))
tracheophytesCleanFixed <- tracheophytesClean

save(tracheophytesCleanFixed, file ="./Intermediate Data/tracheophytesCleanFixed.RData")
```


#2c 
```{r additional corrections, eval = FALSE}
counties <- unique(florida@data$NAME_2)
counties <- factor(counties) 

load("./Intermediate Data/tracheophytesCleanFixed.RData")


#collectors of Anacardiaceae
temp <- subset(tracheophytesCleanFixed, tracheophytesCleanFixed$family == "Anacardiaceae")
temp <- subset(temp, !is.na(temp$correctedCollector))
temp <- subset(temp, !is.na(temp$geo_county))
temp <- subset(temp, !is.na(temp$year))
temp <- subset(temp, !is.na(temp$species))
temp <- subset(temp, temp$species != "")
temp <- subset(temp, temp$month != "")
temp <- subset(temp, !is.na(temp$month))

collectorsAna <- unique(temp$correctedCollector)
collectorsAna <- factor(collectorsAna)

#Tracheophyte occurrences by collectors of Anacardiaceae
trach_by_ana <- subset(tracheophytesCleanFixed, tracheophytesCleanFixed$correctedCollector %in% collectorsAna)
trach_by_ana <- subset(trach_by_ana, !is.na(trach_by_ana$geo_county))
trach_by_ana <- subset(trach_by_ana, !is.na(trach_by_ana$year))
trach_by_ana <- subset(trach_by_ana, !is.na(trach_by_ana$species))
trach_by_ana <- subset(trach_by_ana, trach_by_ana$species != "")
trach_by_ana <- subset(trach_by_ana, trach_by_ana$month != "")
trach_by_ana <- subset(trach_by_ana, !is.na(trach_by_ana$month))

#Check for obviously off dates:
collectors <- NULL
collectors$collector <- unique(trach_by_ana$correctedCollector)
collectors <- as.data.frame(collectors)

for(i in 1:nrow(collectors)) {
  temp <- subset(trach_by_ana, trach_by_ana$correctedCollector == collectors$collector[i])
  collectors$yearMin[i] <- min(temp$correctedYear)
  collectors$yearMax[i] <- max(temp$correctedYear)
  collectors$yearsActive[i] <- length(unique(temp$correctedYear))
  cat("Collector ", i, " of 765 \n")
}

collectors$yearsRange <- as.numeric(collectors$yearMax) - as.numeric(collectors$yearMin)

collectors <- collectors[order(-collectors$yearsRange),]


#Which records are collected by citizen scientists?
#institutionCode's associated with citizen Science
#iNaturalist
#""
#naturgucker
#Bioversity

#basisOfRecord
#"HUMAN_OBSERVATION" (but also pulls up some that are not? because of Natureserve)

#collectionCode 

#Fix any citizen_scientist names that were mistakenly corrected in openRefine

trach_by_ana$citizenScience <- rep(0, nrow(trach_by_ana))
trach_by_ana$correctedCollector <- as.character(trach_by_ana$correctedCollector)
for(i in 1:nrow(trach_by_ana)){
  if(trach_by_ana$basisOfRecord[i] == "HUMAN_OBSERVATION" & trach_by_ana$institutionCode[i] != "NTSRV") {
  trach_by_ana$citizenScience[i] <-1
  trach_by_ana$correctedCollector[i] <- trach_by_ana$recordedBy[i]
  }
  cat("Occurrence record ", i, " of 196611 \n")
}

temp <- subset(trach_by_ana, trach_by_ana$citizenScience == 0)

save(trach_by_ana, file="./Intermediate Data/trach_by_ana_precleanedforyear.RData")







#Check non-citizen-scientist collectors for incorrect collection dates
collectors <- NULL
collectors$collector <- unique(temp$correctedCollector)
collectors <- as.data.frame(collectors)

for(i in 1:nrow(collectors)) {
  temp <- subset(trach_by_ana, trach_by_ana$correctedCollector == collectors$collector[i])
  collectors$yearMin[i] <- min(temp$correctedYear)
  collectors$yearMax[i] <- max(temp$correctedYear)
  collectors$yearsActive[i] <- length(unique(temp$correctedYear))
  cat("Collector ", i, " of 492 \n")
}
collectors$yearRange <- as.numeric(collectors$yearMax) - as.numeric(collectors$yearMin)
collectors <- collectors[order(-collectors$yearRange),]



gbifIDs <- unique(trach_by_ana$gbifID)
correctedDates <- read.csv("./Intermediate Data/correctedDates2.csv")
correctedDates <- subset(correctedDates, correctedDates$gbifID %in% gbifIDs)


for (i in 1:nrow(correctedDates)) {
  if(is.na(correctedDates$correctedYear[i])) {
    trach_by_ana$correctedYear[which(trach_by_ana$gbifID == correctedDates$gbifID[i])] <- NA
  }
  if(correctedDates$correctedYear[i] !="" & !is.na(correctedDates$correctedYear[i])) {
  trach_by_ana$correctedYear[which(trach_by_ana$gbifID == correctedDates$gbifID[i])] <- 
    correctedDates$correctedYear[i] 
  }
   if(is.na(correctedDates$correctedCollector[i])) {
    trach_by_ana$correctedCollector[which(trach_by_ana$gbifID == correctedDates$gbifID[i])] <-     NA
  }
  if(correctedDates$correctedCollector[i] !="" & 
     !is.na(correctedDates$correctedCollector[i])){
        trach_by_ana$correctedCollector[which(trach_by_ana$gbifID == correctedDates$gbifID[i])] <- 
        correctedDates$correctedCollector[i]
  }
  cat("Correcting record ", i, " of ", nrow(correctedDates), "\n")
}


trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "1457749789")] <- 2006
trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "1457830847")] <- 2004
trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "1322791374")] <- 1917
trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "1930250850")] <- 1915
trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "1930009961")] <- 1915
trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "122976884")] <- 1903
trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "123010331")] <- 1913
trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "123003387")] <- 1909


#A. S. Hitchcock (record from 1890 and 1984)
trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "1500137176")] <- 1894
#L. Arnold & E. West
trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "1457754104")] <- 1941
trach_by_ana$correctedYear[which(trach_by_ana$gbifID == "1457813099")] <- 1936

temp <- subset(trach_by_ana, trach_by_ana$family == "Anacardiaceae")
collectorsAna <- unique(temp$correctedCollector)

trach_by_ana <- subset(trach_by_ana, trach_by_ana$correctedCollector %in% collectorsAna)

temp <- subset(trach_by_ana, trach_by_ana$citizenScience == 0)


collectors <- NULL
collectors$collector <- unique(temp$correctedCollector)
collectors <- as.data.frame(collectors)

for(i in 1:nrow(collectors)) {
  temp <- subset(trach_by_ana, trach_by_ana$correctedCollector == collectors$collector[i])
  collectors$yearMin[i] <- min(temp$correctedYear)
  collectors$yearMax[i] <- max(temp$correctedYear)
  collectors$yearsActive[i] <- length(unique(temp$correctedYear))
  cat("Collector ", i, " of 492 \n")
}
collectors$yearRange <- as.numeric(collectors$yearMax) - as.numeric(collectors$yearMin)
collectors <- collectors[order(-collectors$yearRange),]

temp <- subset(trach_by_ana, trach_by_ana$correctedCollector == collectors$collector[34])
sort(unique(temp$year))
unique(temp$recordedBy)
temp2 <- subset(temp, temp$year == 1879)
temp2$gbifID

save(trach_by_ana, file="./Intermediate Data/trach_by_ana_corrected.RData")
nrow(trach_by_ana)
#After this chunk, there are 196,351 records

```


```{r the deduplication, eval = F}


trach_by_ana <- subset(trach_by_ana, as.numeric(trach_by_ana$correctedYear) > 1895)
#Now there are 192,301 records 
deduplicated <- NULL

for (i in counties) { #for each county
  temp <-subset(trach_by_ana, trach_by_ana$geo_county == i)
  collectors_in_county <- unique(temp$correctedCollector)
  for(j in  collectors_in_county) { #for each collector in that county
    temp2 <- subset(temp, temp$correctedCollector == j)
    years_active_in_county <- unique(temp2$correctedYear)
    for(k in years_active_in_county) {
      temp3 <-subset(temp2, temp2$correctedYear == k)
      months_surveyed <- unique(temp3$month)
      for (l in months_surveyed) {
        temp4 <- subset(temp3, temp3$month == l)
        species_surveyed <- unique(temp4$species)
        for (m in species_surveyed) {
          temp5 <- subset(temp4, temp4$species == m) 
          deduplicated <- rbind(deduplicated, temp5[1,])
          cat("County:", i, "Collector:", j, "Year:", k, "Month:", l,
              "\n")
        }
      }
    }
  }
  cat("Finished county", i,"\n")
}

save(deduplicated, file="deduplicated.RData")

#Original deduplicated prior to cleaning years was 166640
#Now with years and collectors cleaned it is 166230

#Check range of date records 

collectors <- NULL
collectors$collector <- unique(deduplicated$correctedCollector)
collectors <- as.data.frame(collectors)

for(i in 1:nrow(collectors)) {
  temp <- subset(deduplicated, deduplicated$correctedCollector == collectors$collector[i])
  collectors$yearMin[i] <- min(temp$correctedYear)
  collectors$yearMax[i] <- max(temp$correctedYear)
  collectors$yearsActive[i] <- length(unique(temp$correctedYear))
  cat("Collector ", i, " of 767 \n")
}


collectors$yearsRange <- as.numeric(collectors$yearMax) - as.numeric(collectors$yearMin)
collectors <- collectors[order(-collectors$yearsRange),]



```


# Create detection history data 

* No filter: All occurrence records 
* Remove singletons: Check whether this is different from removing taxonomic specialists
* Remove taxonomic specialists: Remove records from collectors who only collect one genus (that is not the genus of the focal species).
* Remove non-interested collectors: Remove records from collectors who have never collected the focal species. 
* Month filter: Remove records from months where the focal species was never collected (examine appropriateness of this)

```{r, detection histories and filters, eval=F}

#Load data
load("./florida.RData")
load("./deduplicated.RData")


speciesList <- c("Mangifera indica", "Metopium toxiferum", "Rhus copallina", "Schinus terebinthifolia",
                 "Toxicodendron pubescens", "Toxicodendron radicans", "Toxicodendron vernix")
counties <- florida@data$NAME_2


start_time <- Sys.time()
for(speciesx in speciesList) {

  surv_cov <- data.frame(matrix(ncol = 9, nrow = 0))
  colnames(surv_cov) <- c("county", "year", "month", "collector",
                                "prev_detected", "citizen_scientist", "num_species","num_genera", "h")
  for (i in counties) {
    temp <- subset(deduplicated, deduplicated$geo_county == i)
    temp_collectors <- unique(temp$correctedCollector)
    for(j in  temp_collectors){
      temp2 <- subset(temp, temp$correctedCollector == j)
      temp_years <- sort(unique(temp2$correctedYear))
      for(k in temp_years) {
        temp3 <- subset(temp2, temp2$correctedYear == k)
        temp_months <- sort(unique(temp3$month))
        for (l in temp_months) {
          temp4 <- subset(temp3, temp3$month == l)
          temp_species <- unique(temp4$species)
          temp_genera <- unique(temp4$genus)
          surv_cov_temp <- data.frame(county = i,
                                      year = k,
                                      month = l,
                                      collector = j,
                                      prev_detected = NA,
                                      citizen_scientist = 0,
                                      num_species = length(temp_species),
                                      num_genera = length(temp_genera),
                                      h=0)
          if (speciesx %in% temp_species) {surv_cov_temp$h <- 1}
          if (temp4$basisOfRecord[1] == "HUMAN_OBSERVATION" & temp4$institutionCode[1] != "NTSRV") {
            surv_cov_temp$citizen_scientist <- 1
          }
          surv_cov<- rbind(surv_cov, surv_cov_temp)
          cat("County:", i, "Collector:", j, "Year:", k, "Month:", l, "\n")
        }
      }
    }
  }
  
  surv_cov$county <- factor(surv_cov$county)
  surv_cov$year <- factor(surv_cov$year)
  surv_cov$month  <- factor(surv_cov$month)
  surv_cov$collector <- factor(surv_cov$collector)
  surv_cov$citizen_scientist <- factor(surv_cov$citizen_scientist)
  surv_cov3 <- surv_cov
  save(surv_cov3, file=paste0("./", speciesx, "/surv_cov3.RData"))
}

```


# Gather INFO about collectors of species x #####

```{r, collector covariates}
for (speciesx in speciesList) {
temp <- unique(deduplicated$correctedCollector)

collectors_of_species <- data.frame(collector = temp,
                                totalRecords  = rep(NA, length(temp)),
                                totalSpecies  = rep(NA, length(temp)),
                                totalGenera   = rep(NA, length(temp)),
                                totalFamilies = rep(NA, length(temp)),
                                collectedSpecies = rep(0, length(temp)),
                                yearsActive   = rep(NA, length(temp)))

for (i in 1:length(temp)) {
  temp2 <- subset(deduplicated, deduplicated$correctedCollector == temp[i])
  
  collectors_of_species$totalRecords[i]  <- nrow(temp2)
  collectors_of_species$totalSpecies[i]  <- length(unique(temp2$species))
  collectors_of_species$totalGenera[i]   <- length(unique(temp2$genus))
  collectors_of_species$totalFamilies[i] <- length(unique(temp2$family))
  collectors_of_species$yearsActive[i]   <- 
    length(unique(temp2$correctedYear))
  if(speciesx %in% temp2$species) {collectors_of_species$collectedSpecies[i] <- 1}
  cat("Collector ", i, " of ", length(temp), "\n")
}

save(collectors_of_species,
     file=paste0("./", speciesx, "/collectors_of_species.RData"))
}

```



# FILTRATION #####
* No filter: All occurrence records 
* Remove singletons: Check whether this is different from removing taxonomic specialists
* Remove taxonomic specialists: Remove records from collectors who only collect one genus (that is not the genus of the focal species).
* Remove non-interested collectors: Remove records from collectors who have never collected the focal species. 
* Month filter: Remove records from months where the focal species was never collected (examine appropriateness of this)
```{r, filtration, eval=F}

for (speciesx in speciesList) {
  load(paste0("./", speciesx, "/surv_cov3.RData"))
  load(paste0("./", speciesx, "/collectors_of_species.RData"))
  
  #Filter out surveys where collectors only collected one species (unless it's a focal species)
  not_singletons <- subset(collectors_of_species, collectors_of_species$totalSpecies>1)
  not_single_collectors <- unique(not_singletons$collector)
  surv_cov1 <- subset(surv_cov, surv_cov$h==1 |(surv_cov$collector %in% not_single_collectors))
  save(surv_cov1,
       file = paste0("./", speciesx, "/surv_cov1.RData"))
  
   #Genus filter (Remove taxonomic specialists):
  temp <- subset(collectors_of_species, collectors_of_species$totalGenera >1 |
                   collectors_of_species$collectedSpecies ==1)
  temp2 <- unique(temp$collector)
  surv_cov_genus <- subset(surv_cov1, surv_cov1$collector %in% temp2)
  save(surv_cov_genus,
       file = paste0("./", speciesx, "/surv_cov_genus.RData"))
  
  #Species filter: 
  temp <- subset(deduplicated, deduplicated$species == speciesx)
  temp2 <- unique(temp$correctedCollector) 
  surv_cov_species <- subset(surv_cov1, surv_cov1$collector %in% temp2)
  save(surv_cov_species,
       file = paste0("./", speciesx, "/surv_cov_species.RData"), version=2)

  #Month filter
  temp <- subset(deduplicated, deduplicated$species == speciesx)
  temp2 <- unique(temp$month)
  temp3 <- subset(collectors_of_species, collectors_of_species$collectedSpecies == 1)$collector
  
  surv_cov_month <- subset(surv_cov1, surv_cov1$month %in% temp2 | surv_cov1$collector %in% temp3)
  save(surv_cov_month,
       file = paste0("./", speciesx, "/surv_cov_month.RData"))
}

total_time <- Sys.time() - start_time
```



# FILL OUT PREV_DETECTED AND REFACTOR #####

```{r fix some issues}

for(speciesx in speciesList) {
  
  #Populate previously detected variable for surv_cov model 
  load(paste0("./", speciesx, "/surv_cov3.RData"))
  surv_cov <- surv_cov3
  rm(surv_cov3)
  surv_cov$index <- 1:length(surv_cov$county)
  collectors <- unique(surv_cov$collector)
  for (i in collectors) {
    temp <- subset(surv_cov, surv_cov$collector == i)
    counties <- unique(temp$county)
    for (j in counties){
      temp2 <- subset(temp, temp$county == j)
      if (sum(temp2$h) == 0) {#collector has not seen this species in this county {
        indices <- unique(temp2$index)
        surv_cov$prev_detected[indices] <- 0
      }
      if (sum(temp2$h) > 0) { #collector HAS seen this species in this county}
        indices <- unique(temp2$index)
        surv_cov$prev_detected[indices[1]] <- 0
        if(length(indices)>1) {surv_cov$prev_detected[indices[2:length(indices)]] <- 1}
      }
    }
  }
  
  surv_cov$year <- factor(surv_cov$year)
  surv_cov$collector <- factor(surv_cov$collector)
  surv_cov3<-surv_cov
  save(surv_cov3, file=paste0("./", speciesx, "/surv_cov3.RData"))
  
}
  #######Not doing things underneath this
  #Populate previously detected variable for surv_cov_species model:
  load(paste0("./", speciesx, "/surv_cov_species.RData"))
  surv_cov_species$index <- 1:length(surv_cov_species$county)
  collectors <- unique(surv_cov_species$collector)
  for (i in collectors) {
    temp <- subset(surv_cov_species, surv_cov_species$collector == i)
    counties <- unique(temp$county)
    for (j in counties){
      temp2 <- subset(temp, temp$county == j)
      if (sum(temp2$h) == 0) {#collector has not seen this species in this county {
        indices <- unique(temp2$index)
        surv_cov_species$prev_detected[indices] <- 0
      }
      if (sum(temp2$h) > 0) { #collector HAS seen this species in this county}
        indices <- unique(temp2$index)
        surv_cov_species$prev_detected[indices[1]] <- 0
        if(length(indices)>1) {surv_cov_species$prev_detected[indices[2:length(indices)]] <- 1}
      }
    }
  }
  surv_cov_species$year <- factor(surv_cov_species$year)
  surv_cov_species$collector <- factor(surv_cov_species$collector)
  save(surv_cov_species, file=paste0("./", speciesx, "/surv_cov_species.RData"))
  
  #Populate previously detected variable for surv_cov_genus model:
  load(paste0("./", speciesx, "/surv_cov_genus.RData"))
  surv_cov_genus$index <- 1:length(surv_cov_genus$county)
  collectors <- unique(surv_cov_genus$collector)
  for (i in collectors) {
    temp <- subset(surv_cov_genus, surv_cov_genus$collector == i)
    counties <- unique(temp$county)
    for (j in counties){
      temp2 <- subset(temp, temp$county == j)
      if (sum(temp2$h) == 0) {#collector has not seen this species in this county {
        indices <- unique(temp2$index)
        surv_cov_genus$prev_detected[indices] <- 0
      }
      if (sum(temp2$h) > 0) { #collector HAS seen this species in this county}
        indices <- unique(temp2$index)
        surv_cov_genus$prev_detected[indices[1]] <- 0
        if(length(indices)>1) {surv_cov_genus$prev_detected[indices[2:length(indices)]] <- 1}
      }
    }
  }
  surv_cov_genus$year <- factor(surv_cov_genus$year)
  surv_cov_genus$collector <- factor(surv_cov_genus$collector)
  save(surv_cov_genus, file=paste0("./", speciesx, "/surv_cov_genus.RData"))
  
  #Populate previously detected variable for surv_cov_month model:
  load(paste0("./", speciesx, "/surv_cov_month.RData"))
  surv_cov_month$index <- 1:length(surv_cov_month$county)
  collectors <- unique(surv_cov_month$collector)
  for (i in collectors) {
    temp <- subset(surv_cov_month, surv_cov_month$collector == i)
    counties <- unique(temp$county)
    for (j in counties){
      temp2 <- subset(temp, temp$county == j)
      if (sum(temp2$h) == 0) {#collector has not seen this species in this county {
        indices <- unique(temp2$index)
        surv_cov_month$prev_detected[indices] <- 0
      }
      if (sum(temp2$h) > 0) { #collector HAS seen this species in this county}
        indices <- unique(temp2$index)
        surv_cov_month$prev_detected[indices[1]] <- 0
        if(length(indices)>1) {surv_cov_month$prev_detected[indices[2:length(indices)]] <- 1}
      }
    }
  }
  surv_cov_month$year <- factor(surv_cov_month$year)
  surv_cov_month$collector <- factor(surv_cov_month$collector)
  save(surv_cov_month, file=paste0("./", speciesx, "/surv_cov_month.RData"))
  
  #Populate previously detected variable for surv_cov1 model:
  load(paste0("./", speciesx, "/surv_cov1.RData"))
  surv_cov1$index <- 1:length(surv_cov1$county)
  collectors <- unique(surv_cov1$collector)
  for (i in collectors) {
    temp <- subset(surv_cov1, surv_cov1$collector == i)
    counties <- unique(temp$county)
    for (j in counties){
      temp2 <- subset(temp, temp$county == j)
      if (sum(temp2$h) == 0) {#collector has not seen this species in this county {
        indices <- unique(temp2$index)
        surv_cov1$prev_detected[indices] <- 0
      }
      if (sum(temp2$h) > 0) { #collector HAS seen this species in this county}
        indices <- unique(temp2$index)
        surv_cov1$prev_detected[indices[1]] <- 0
        if(length(indices)>1) {surv_cov1$prev_detected[indices[2:length(indices)]] <- 1}
      }
    }
  }
  
  surv_cov1$year <- factor(surv_cov1$year)
  surv_cov1$collector <- factor(surv_cov1$collector)
  save(surv_cov1, file=paste0("./", speciesx, "/surv_cov1.RData"))


}


```

# PRISM DATA

```{r prism}
#1) Convert the original files downloaded from PRISM into the appropriate file type: #####

#Only have to run this once

var <- 'tmin'
for (year in 2016:2018) {
  
  for (mo in 1:12) {
  
    r <- raster(paste0(workDir, 'PRISM/2 arcmin/ORIGINALS/', var, '/', year, '/', 'PRISM_', var, '_stable_4kmM3_', year, prefix(mo, 2), '_bil.bil'))
 
    writeRaster(r, paste0(workDir,var, '/', 'PRISM_',var, '_stable_4kmM2_', year, prefix(mo, 2), '_bil'), format='GTiff', datatype='INT4S', overwrite=TRUE)
    
  }
  
}

#2) Stack and calculate tmin_min annual temperature raster #####
var <- 'tmin'
fun <- 'min'



albersNA <- '+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs'
wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'

dirCreate <- function(...) base::dir.create(path=paste0(...), recursive=TRUE, showWarnings=FALSE)
say <- function(..., pre=0, post=1) { cat(rep('\n', pre), ..., rep('\n', post)); flush.console() }

prefix <- function(x, len, pre='0') {

	# x		value to add leading characters (will be coerced to character class)
	# len	desired number of characters in x... will not be made shorter if x is longer than this
	# pre	value to pre-pend to x, will be repeated until nchar(x)==len
	
	x <- as.character(x)
	size <- nchar(x)
	if (nchar(x) < len) {
		addTo <- paste(rep(pre, each=len - size), collapse='')
		x <- paste0(addTo, x)
	}
	return(x)
}

var <- 'tmin'



fun <- 'min'


for (countYear in 1895:2018) {

	say(countYear)

	yearStack <- stack(list.files(paste(workDir, 'PRISM/2 arcmin/', var, sep=''),
	                              pattern=paste('PRISM_', var, '_stable_4kmM2_', countYear, sep=''), countYear, full.names=T))

	newRast <- if (fun=='mean') {
		mean(yearStack)
	} else if (fun=='max') {
		max(yearStack)
	} else if (fun=='min') {
		min(yearStack)
	} else if (fun=='sum') {
		sum(yearStack)
	}

	newRast <- setMinMax(newRast)
	projection(newRast) <- projection(yearStack)
	
	writeRaster(newRast, paste(workDir, 'PRISM/2 arcmin/', var, '/', var, '_annual_', fun, '/', var, '_', fun, '_', countYear, sep=''), format='GTiff', overwrite=T)
	
	rm(yearStack, newRast)
	gc()
	
}


#3) Clip rasters to florida #####



load("./florida.RData")
#rast <- crop(mat, florida)

#for (i in 1895:2015) {
 for(i in 1895:2018) { 
  
  # get the world raster
  rast <- raster(paste0('./PRISM/2 arcmin/tmin/tmin_annual_min/tmin_min_',
                         i, '.tif'))
  
  # crop the raster and update metadata
  rast <- crop(rast, florida)
  rast <- setMinMax(rast)
  projection(rast) <-
    '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
  
  # save raster
  writeRaster(
    rast,
    paste0('./PRISM/2 arcmin/florida/tmin_min_',
           i), format='GTiff', datatype='INT2S',
    overwrite=TRUE
  )
}


tmin_min_avg <- matrix(nrow=67, ncol=124)
#years <- 1895:2015
years <- 1895:2018

#Originally did 1895:2016. Now need to do 2016:2018
#temp <- rep(NA, 67)
#tmin_min_avg <- cbind(tmin_min_avg, thing, thing, thing)
for(i in 1:124) {
  print(paste('Calculating average county tmin for year: ', years[i]))
  rast <- raster(paste0('./PRISM/2 arcmin/florida/tmin_min_',
                        years[i], '.tif'))
  
  temp <- extract(rast, florida)
  thing <-rep(NA, 67)
  for (j in 1:67) {
    thing[j] <- mean(temp[[j]], na.rm=T)
  }
  tmin_min_avg[,i] <- thing
}

save(tmin_min_avg, file="./PRISM/2 arcmin/florida/tmin_min_avg.RData")


#Calculate average for each county over all 124 years. Composite_average_tmin_min represents the average of the 
# minimum minimum temperature in a county across all 121 years from 1895-2015. 

composite_average_tmin_min <- rep(NA, 67)
for (i in 1:67) {
  composite_average_tmin_min[i] <- mean(tmin_min_avg[i,])
}

save(composite_average_tmin_min, file="./composite_average_tmin_min.RData")


```
# Assemble the covariates associated with the counties

```{r county covariates, eval=F }
unit_cov <- data.frame(
  county = rep(NA, 67),
  pop2010 = rep(NA, 67),
  tmin_min_avg = rep(NA, 67),
  area = rep(NA, 67)
)

#Human Population Density
load("census_2010.RData")
census_2010 <- census_2010[order(census_2010$county),]

unit_cov$county <- florida$NAME_2
unit_cov$pop2010 <- census_2010$PCT001001

#Average Minimum Temperature 
load("./composite_average_tmin_min.RData")
unit_cov$tmin_min_avg <- composite_average_tmin_min

#County area

florida2 <- spTransform(florida, CRS("+init=epsg:26978")) 
unit_cov$area <- gArea(florida2, byid=T)
save(unit_cov, file="./unit_cov.RData")


unit_cov3 <- unit_cov
# Standardize the unit_cov statistics: 
hist(unit_cov$area)
hist(unit_cov$pop2010)
#Both of these seem to have extreme outliers, so first I will log transform population size and area before z-transforming

unit_cov2 <- unit_cov


unit_cov3$pop2010 <- log(unit_cov$pop2010)
unit_cov3$area    <- log(unit_cov$area)
unit_cov3$tmin_min_avg <- log(unit_cov$tmin_min_avg)

hist(unit_cov2$pop2010)
hist(unit_cov2$area)


mean(unit_cov2$pop2010)
sd(unit_cov2$pop2010)

unit_cov2$pop2010 <- standardize(unit_cov3$pop2010)
unit_cov2$tmin_min_avg <- standardize(unit_cov3$tmin_min_avg)
unit_cov2$area <- standardize(unit_cov3$area)


save(unit_cov2, file="./unit_cov2.RData")
```
